HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2023-02-02 tmlr_submission/hermes_paper_submission/hermes_paper.tex
Minor updates on Section 3

In theoretical fashion dynamics \citep{rogers1962}, different categories of adopters follow a trend in succession, resulting in several adoption waves.
So as too catch the early signal of the emergence of a trend, 6000 social media influencers were selectioned by hand by fashion experts. Aggregating them, a specific ``fashion-oriented`` panel is created. With the same methodology as for the main panel described in Section~\ref{sec:dataset:a} and Section~\ref{sec:dataset:b}, a normalized time series representing each fashion trend on this specific population is created. We named \textit{fashion-forwards} this weak signal.  For all fashion sequence $\{y^i_t\}_{1 \leq t \leq T}$, let $\{\ts^{f,i}_t\}_{1 \leq t \leq T}$ be the normalized sequence representing the behaviours of influencers regarding the fashion trend $i$. As we want to detect shifts between the main signal and the fashion forward signal, the following input is computed for the hybrid model: for all $t \in \{1,\ldots,T\}$ and any fashion trend $i$,
$$
\ws^{f,i}_{t} = \frac{\ts_t^{f,i}}{\ts_t^{f,i}+\ts_t^{i}}\,.
$$
where T denotes the number of available time steps. Values close to 0.5 indicate a similar behaviour between the influencers panel and the general panel. For instance, an  emerging fashion shoes trend with its \textit{fashion-forwards} weak signal is represented in Figure~\ref{fig:oneemergingtrend}.

@@

%In the specific application of fashion trends prediction, two main forms of standard weaknesses were detected and defined for univariate parametric forecasts. The first one is a weakness of definition, when the model is not well defined for its forecasting task. For instance, an additive model provides good predictions on time series with additive seasonality but fails at handling multiplicative seasonalities. The second form, harder to detect and correct, is the weakness of information. 
For non stationary time series, huge changes of behaviours are not always predictable using the past of the sequence. In some cases, these changes depend on external variables not considered by univariate parametric models. The difficulty is that the exact influence of external variables on the main signal is mostly unknown. This motivates the introduction of a global RNN trained on all time series and able to consider and leverage external signals. First forecast example of our new hybrid model is given in Figure~\ref{fig:introexamples}.

@@

The second part of the model is a global RNN, trained on all the $N$ sequences to correct the weaknesses of the first per-time-series parametric models. This task requires a thorough data pre-processing as recurrent neural networks training is highly sensitive to the scale of the data and requires well-designed inputs. Since no assumption about the scale of our time series was made, inputs require a careful normalization before being fed to the RNN.

Let $\window \in \mathbb{N}$ be the window size, usually this window is proportional to the forecast horizon $\window \propto \lag$. The RNN input is defined as the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnninput^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$: for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$,
$$
\rnninput^{n,T}_{T-w+i|T} := \frac{\ts^n_{T-w+i} -\tspred^{pred,n}_{T+k|T}}{\meants^n_T}\,,\quad \meants^n_T = \frac{1}{w}\sum_{i = 1}^{w}\ts^n_{T-w+i}\,.
$$
where $k = i - h\lfloor i/h \rfloor$ with $\lfloor. \rfloor$ the floor function. With the numerator part $\ts^n_{T-w+i} -\tspred^{pred,n}_{T+k|T}$, the per-time-series prediction is included in the RNN input and all the fundamental patterns already learned by this first predictor are removed from the time series. Then the denominator $\meants^n_T$ is use to rescaled all input at the same level as the time series can have different scales. Another option could have been to divide directly  $\ts^n_{T-w+i}$ by $\tspred^{pred,n}_{T+k|T}$ but with time series hitting 0, this option is not valid. Let $\rnn(.;\rnnparam)$ be the recurrent neural network model where $\rnnparam$ are  unknown parameters. Given the RNN input sequences $\{\rnnwindow^n_T\}_{1\leqslant n \leqslant N}$ and the global RNN estimated parameters $\rnnparam$, the error-corrector predictions $\{\tspred^{corr,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{corr,n}_{T+i|T} = \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\ \,.
$$
Our hybrid model forecast is, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,\lag\}$,
\begin{align}
\label{eq:nows:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T} \\
&= \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}

@@

%Using well-fitted time-series-specific parametric models, the new hybrid network corrects the first form of weakness and provides appealing performance on the fashion dataset, see Table~\ref{tab:metricresults}. 
%Then, to correct the second form of weakness, 
In addition to the $N$ target time series, $K \times N$ external sequences indexed from $0$ to $T$ are considered. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ associated with the sequence $\fullts^n$. Let  $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ be all the values of the weak signals associated with the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the sequence. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, a new input for the RNN is defined:   
$$
\fullconcatinput^n_T = \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
$$
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ the final prediction becomes:

\begin{align}
\label{eq:withws:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T}\\
& = \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\fullconcatinput^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}
An illustration of the proposed  model is displayed in Figure~\ref{fig:architecture}.

>>>>

In theoretical fashion dynamics \citep{rogers1962}, different categories of adopters follow a trend in succession, resulting in several adoption waves.
So as to catch the early signal of the emergence of a trend, 6000 social media influencers were selectioned by hand by fashion experts. Aggregating them, a specific ``fashion-oriented`` panel is created. With the same methodology as for the main panel described in Section~\ref{sec:dataset:a} and Section~\ref{sec:dataset:b}, a normalized time series representing each fashion trend on this specific population is created. We named \textit{fashion-forwards} this weak signal.  For all fashion sequence $\{y^i_t\}_{1 \leq t \leq T}$, let $\{\ts^{f,i}_t\}_{1 \leq t \leq T}$ be the normalized sequence representing the behaviours of influencers regarding the fashion trend $i$. As we want to detect shifts between the main signal and the fashion forward signal, the following input is computed for the hybrid model: for all $t \in \{1,\ldots,T\}$ and any fashion trend $i$,
$$
\ws^{f,i}_{t} = \frac{\ts_t^{f,i}}{\ts_t^{f,i}+\ts_t^{i}}\,.
$$
where T denotes the number of available time steps. Values close to 0.5 indicate a similar behaviour between the influencers panel and the general panel. For instance, an  emerging fashion shoes trend with its \textit{fashion-forwards} weak signal is represented in Figure~\ref{fig:oneemergingtrend}.

@@

%In the specific application of fashion trends prediction, two main forms of standard weaknesses were detected and defined for univariate parametric forecasts. The first one is a weakness of definition, when the model is not well defined for its forecasting task. For instance, an additive model provides good predictions on time series with additive seasonality but fails at handling multiplicative seasonalities. The second form, harder to detect and correct, is the weakness of information. 
For non stationary time series, huge changes of behaviours are not always predictable using the past of the sequence. In some cases, these changes depend on external variables not considered by univariate parametric models. The difficulty is that the exact influence of external variables on the main signal is mostly unknown. This motivates the introduction of a global RNN trained on all time series and able to consider and leverage external signals.

@@

The second part of the model is a global RNN, trained on all the $N$ sequences to correct the weaknesses of the first per-time-series parametric models. This task requires a thorough data pre-processing as recurrent neural networks training is highly sensitive to the scale of the data and requires well-designed inputs.

Let $\window \in \mathbb{N}$ be the window size, usually this window is proportional to the forecast horizon $\window \propto \lag$. The RNN input is defined as the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnninput^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$: for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$,
$$
\rnninput^{n,T}_{T-w+i|T} := \frac{\ts^n_{T-w+i} -\tspred^{pred,n}_{T+k|T}}{\meants^n_T}\,,\quad \meants^n_T = \frac{1}{w}\sum_{i = 1}^{w}\ts^n_{T-w+i}\,.
$$
where $k = i - h\lfloor i/h \rfloor$ with $\lfloor. \rfloor$ the floor function. With the numerator part $\ts^n_{T-w+i} -\tspred^{pred,n}_{T+k|T}$, the per-time-series prediction is included in the RNN input and all the fundamental patterns already learned by this first predictor are removed from the time series. Then the denominator $\meants^n_T$ is use to rescaled all input at the same level as the time series can have different scales. Another option could have been to divide directly  $\ts^n_{T-w+i}$ by $\tspred^{pred,n}_{T+k|T}$ but with time series hitting 0, this option is not valid. Let $\rnn(.;\rnnparam)$ be the recurrent neural network model where $\rnnparam$ are  unknown parameters. Given the RNN input sequences $\{\rnnwindow^n_T\}_{1\leqslant n \leqslant N}$ and the global RNN estimated parameters $\rnnparam$, the error-corrector predictions $\{\tspred^{corr,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{corr,n}_{T+i|T} = \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\ \,.
$$
Thus, if no external signals are available, the final hermes  forecast is, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,\lag\}$,
\begin{align}
\label{eq:nows:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T} \\
&= \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}

@@

%Using well-fitted time-series-specific parametric models, the new hybrid network corrects the first form of weakness and provides appealing performance on the fashion dataset, see Table~\ref{tab:metricresults}. 
%Then, to correct the second form of weakness, 
In addition to the $N$ target time series, $K \times N$ external sequences indexed from $0$ to $T$ are now considered. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ associated with the sequence $\fullts^n$. Let  $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ be all the values of the external signals. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the external sequences. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, a new input for the RNN is defined:   
$$
\fullconcatinput^n_T = \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
$$
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ the final prediction becomes:

\begin{align}
\label{eq:withws:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T}\\
& = \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\fullconcatinput^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}
An illustration of the proposed  model is displayed in Figure~\ref{fig:architecture} and a first forecast example is given in Figure~\ref{fig:introexamples}..
